Depression and suicidality as costly signals of need

### 2021-04-08 Paper.rmd
Demographic models now US only (to reduce confounds).
Updated some packages.

tar_load(
    c(
    'd',
    'signalingdata2018b',
    'signaling2020',
    'pilot_results',
    'power_curve',
    'power_curve2',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'ecdf_plot',
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'plot_raw_belief_action2d',
    'med_prereg',
    'med1',
    'med1b',
    'med4',
    'med4b',
    'm1_boot',
    'm3_boot',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
)

@@

We therefore fit models of our two outcome variables (which we rescaled to the continuous interval [0, 1]) in four ways. First, we fit the preregistered OLS models (reported in the SI). Second, we fit fractional regression models, introduced in a seminal paper by @papke1996 for continuous outcomes on the interval [0, 1], which, in brief, involve fitting a quasi-likelihood version of logistic regression with robust standard errors. This method is implemented in the R `frm` package [@ramalho2019]. Third, we fit GLM logistic regression models, computing 95% CI's via bootstrapping, using the R `glmmboot` package created to model continuous outcomes on [0, 1] [@humphrey2020]. Finally, we fit GLM models with the quasibinomial family and logit link using the `glm` function in R stats library, which can accept continous outcomes on [0, 1]. The estimated coefficients, standard errors, and confidence intervals produced by the latter three methods were *very* similar, and often identical. We therefore proceeded to fit and report models using `glm` with the quasibinomilal family and logit link because model objects produced by this function are compatible with numerous other R packages for estimating and plotting effects.

Although two of our four preregistered OLS regression models included an interaction term, the rationale for it was less convincing for quasibinomial models using a logit link, which transforms [0, 1] to (-$\infty$, $\infty$). We therefore only included it if there was clear evidence of superior fit. Our intervention was *signal*, coded as an ordinal variable with the following preregistered rank order: verbal request, crying, mild depression, depression, suicide attempt. We used the default `glm` settings, which code such variables as orthogonal polynomial contrasts. Our primary confirmatory regression models were as follows (where the signal subscripts indicate polynomial degree):

@@

To test our hypothesis that the likelihood of acting to help the victim would be largely mediated by a signal's positive impact on the participant's beleif in the victim's need, we used the `mediation` package [@tingley2014] to fit a mediation model for *Depression* vs. *Verbal request*. The mediation and outcome models, respectively, were:

@@

vars <- c(
  Age = 'Age (years)',
  Income = 'Income (USD)',
  minutes = 'Time to complete (minutes)',
  T1Belief = 'T1 Belief',
  T1Action = 'T1 Action',
  T1Divide = 'T1 Division',
  T1LowMood = 'T1 Low mood',
  T1Manipulative = 'T1 Manipulative',
  T2Belief = 'T2 Belief',
  T2Action = 'T2 Action',
  T2Divide = 'T2 Division',
  T2LowMood = 'T2 Low mood',
  T2Manipulative = 'T2 Manipulative',
  T3Action = 'T3 Action'
)
kable(custom.summarize(d, vars), caption = "Summary statistics for study variables. Original 0-100 values scaled to 0-1. Indian participants reported income in rupees, which we converted to USD at the current exchange rate (1 rupee = 0.014 USD).") %>% kable_styling()
```

@@

suicide_belief_eff <- signif(models$Margins$m1$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

suicide_action_eff <- signif(models$Margins$m3$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

@@

As predicted, there was a strong statistically significant positive causal effect of the ordinal signal on T2 Belief and Action (controlling for T1 Belief and Action, respectively), with the effect of each signal increasing in the predicted rank order. The marginal between-subjects effect of the high cost *Suicide attempt* vs. low cost *Verbal request* signal on T2 *Belief* (averaging over all four vignettes and all values of T1 *Belief*) was an increase of `r fmt_margins(models$Margins$m1, 'signalSuicide attempt')` points on the original 0-100 point scale. The equivalent increase for T2 *Action* was `r fmt_margins(models$Margins$m3, 'signalSuicide attempt')` points. Given that the standard deviations of T2 *Belief* and *Action* are `r signif(sd(100*d$T2Belief), 2)` and `r signif(sd(100*d$T2Action), 2)`, respectively, these increases represent `r suicide_belief_eff` and `r suicide_action_eff` SD increases, respectively.

There was modest evidence for a nonlinear effect of the signals on *Belief* and *Action*, which, based on visual inspection, was probably due to the *Mild depression* and *Depression* signals both having similar large effects compared to *Crying* and *Verbal request*. See Figure \@ref(fig:beliefeffect) and model m1 in Tables \@ref(tab:regressiontables) and \@ref(tab:anovatables).

@@

We note the following patterns, which we will return to in the Discussion section. First, as we intended, the distributions of *Belief* and *Action* in the *Verbal request* condition were very similar to their distributions at T1 baseline across vignettes. In the "Basketball coach" and "Brother-in-law" vignettes, though, their distributions in the *Verbal request* condition were shifted to somewhat lower values relative to baseline (i.e., the verbal request slightly *reduced* belief and action in those vignettes). Second, in the "Romantic partner" and "Brother-in-law" vignettes, the effect of *Crying* differed little from *Verbal request*, but in the "Basketball coach" vignette, it differed little from the *Depression* and *Suicide attempt* signals. Third, the effect of the *Suicide attempt* signal on *Belief* was similar to that of *Depression* across all vignettes, but had a noticeably greater effect on *Action* in the "Romantic partner" and "Brother-in-law" vignettes. Fourth, there was little difference between the effect of the *Mild depression* vs. *Depression* signals. See Figures \@ref(fig:beliefvignette) and \@ref(fig:ecdf).

@@

Overall, women participants were more likely to believe and help the victim than men, and younger participants were more likely to believe and help the victim than older participants. The caveats are that the sex differences in *Belief* only emerged for the costly signals, the sex differences in *Action* were quite small (but applied across all signals), and the sex and age effects in the Indian sample were non-existent or slightly reversed. We found no significant effects of *Income* on *Belief* or *Action* (results not reported). See Figures \@ref(fig:demographicsignal), \@ref(fig:demovignettebelief), and \@ref(fig:demovignetteaction).

@@

(effectsPlots$m24plotAge + ylim(0, 1) + effectsPlots$m25plotAge + ylim(0, 1))  / (effectsPlots$m24plotSex + effectsPlots$m25plotSex + plot_layout(widths = c(5,1)))

@@

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Belief by vignette, controlling for T1 Belief and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
(effectsPlots$m18plot / effectsPlots$m13plot)
```

```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Action by vignette, controlling for T1 Action and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots$m19plot / effectsPlots$m15plot
```

>>>>

tar_load(
    c(
    'd',
    'signalingdata2018b',
    'signaling2020',
    'pilot_results',
    'power_curve',
    'power_curve2',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'effectsPlots_US',
    'effectsPlots_India',
    'ecdf_plot',
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'plot_raw_belief_action2d',
    'med_prereg',
    'med1',
    'med1b',
    'med4',
    'med4b',
    'm1_boot',
    'm3_boot',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
)

@@

We therefore fit models of our two outcome variables (which we rescaled to the continuous interval [0, 1]) in four ways. First, we fit the preregistered OLS models (reported in the SI). Second, we fit fractional regression models, introduced in a seminal paper by @papke1996 for continuous outcomes on the interval [0, 1], which, in brief, involve fitting a quasi-likelihood version of logistic regression with robust standard errors. This method is implemented in the R `frm` package [@ramalho2019]. Third, we fit GLM logistic regression models, computing 95% CI's via bootstrapping, using the R `glmmboot` package created to model continuous outcomes on [0, 1] [@humphrey2020]. Finally, we fit GLM models with the quasibinomial family and logit link using the `glm` function in R stats library, which can accept continuous outcomes on [0, 1]. The estimated coefficients, standard errors, and confidence intervals produced by the latter three methods were *very* similar, and often identical. We therefore proceeded to fit and report models using `glm` with the quasibinomial family and logit link because model objects produced by this function are compatible with numerous other R packages for estimating and plotting effects.

Although two of our four preregistered OLS regression models included an interaction term, the rationale for it was less convincing for quasibinomial models using a logit link, which maps [0, 1] to (-$\infty$, $\infty$). The interaction would also increase the number of parameters to estimate, increasing the risk of overfitting. We therefore did not include it in the GLM models (but we still report OLS models with the interaction term).

Our intervention was *signal*, an ordinal variable with the following preregistered rank order: verbal request, crying, mild depression, depression, suicide attempt. We used the default `glm` settings, which code such variables as orthogonal polynomial contrasts. Our primary confirmatory regression models were as follows (where the signal subscripts indicate polynomial degree):

@@

To test our hypothesis that the likelihood of acting to help the victim would be largely mediated by a signal's positive impact on the participant's belief in the victim's need, we used the `mediation` package [@tingley2014] to fit a mediation model for *Depression* vs. *Verbal request*. The mediation and outcome models, respectively, were:

@@

vars <- c(
  Age = 'Age (years)',
  Income = 'Income (USD)',
  years_education = 'Education (years)',
  minutes = 'Time to complete (minutes)',
  T1Belief = 'T1 Belief',
  T1Action = 'T1 Action',
  T1Divide = 'T1 Division',
  T1LowMood = 'T1 Low mood',
  T1Manipulative = 'T1 Manipulative',
  T2Belief = 'T2 Belief',
  T2Action = 'T2 Action',
  T2Divide = 'T2 Division',
  T2LowMood = 'T2 Low mood',
  T2Manipulative = 'T2 Manipulative',
  T3Action = 'T3 Action'
)
kable(custom.summarize(d, vars), caption = "Summary statistics for study variables. Original 0-100 values scaled to 0-1. Indian participants reported income in rupees, which we converted to USD at the current exchange rate (1 rupee = 0.014 USD).") %>% kable_styling()
```

@@

```{r}
sm1 <- summary(models$Margins$m1)
suicide_belief_eff <- signif(sm1$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

sm3 <- summary(models$Margins$m3)
suicide_action_eff <- signif(sm3$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

@@

As predicted, there was a strong statistically significant positive causal effect of the ordinal signal on T2 Belief and Action (controlling for T1 Belief and Action, respectively), with the effect of each signal increasing in the predicted rank order. See Figure \@ref(fig:beliefeffect) and models m1 and m3 in Tables \@ref(tab:regressiontables) and \@ref(tab:anovatables).

The marginal between-subjects effect of the high cost *Suicide attempt* vs. low cost *Verbal request* control on T2 *Belief* (averaging over all four vignettes and all values of T1 *Belief*) was an increase of `r fmt_margins(models$Margins$m1, 'signalSuicide attempt')` points on the original 0-100 point scale. The equivalent increase for T2 *Action* was `r fmt_margins(models$Margins$m3, 'signalSuicide attempt')` points. Given that the standard deviations of T2 *Belief* and *Action* are `r signif(sd(100*d$T2Belief), 2)` and `r signif(sd(100*d$T2Action), 2)`, respectively, these represent increases of `r suicide_belief_eff` and `r suicide_action_eff` SD's, respectively.

@@

We note the following patterns, which we will return to in the Discussion section. First, as we intended, the distributions of *Belief* and *Action* in the *Verbal request* control condition were very similar to their distributions at T1 baseline across vignettes. In the "Basketball coach" and "Brother-in-law" vignettes, though, their distributions in the *Verbal request* condition were shifted to somewhat lower values relative to baseline (i.e., the verbal request slightly *reduced* belief and action in those vignettes). Second, in the "Romantic partner" and "Brother-in-law" vignettes, the effect of *Crying* differed little from *Verbal request*, but in the "Basketball coach" vignette, it differed little from the *Depression* and *Suicide attempt* signals. Third, the effect of the *Suicide attempt* signal on *Belief* was similar to that of *Depression* across all vignettes, but had a noticeably greater effect on *Action* in the "Romantic partner" and "Brother-in-law" vignettes. Fourth, there was little difference between the effect of the *Mild depression* vs. *Depression* signals. See Figures \@ref(fig:beliefvignette) and \@ref(fig:ecdf).

@@

The demographic variables were strongly confounded with the Indian sample (younger, lower income, more years of education) vs. the US sample (older, higher income, fewer years of education), which, in turn, were confounded with vignette (responses in the Thwarted marriage vignette differed substantially from those in the other vignettes). We therefore restricted our exploration of the sociodemographic variables to the US sample.

Overall, women participants were more likely to believe and help the victim than men, and younger participants were more likely to believe and help the victim than older participants. The caveats are that the sex differences in *Belief* only emerged for the costly signals, the sex differences in *Action* were quite small (but applied across all signals). We found no significant effects of *Income* or *Education* on *Belief* or *Action* (results not reported). See Figures \@ref(fig:demographicsignal), \@ref(fig:demovignettebelief), and \@ref(fig:demovignetteaction).

@@

(effectsPlots_US$m24plotAge + ylim(0, 1) + effectsPlots_US$m25plotAge + ylim(0, 1))  / (effectsPlots_US$m24plotSex + effectsPlots_US$m25plotSex + plot_layout(widths = c(5,1)))

@@

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Belief by vignette, controlling for T1 Belief and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
(effectsPlots_US$m18plot / effectsPlots_US$m13plot)
```

```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Action by vignette, controlling for T1 Action and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots_US$m19plot / effectsPlots_US$m15plot
```
